### 2018-10-25 pkg/model4you/inst/JORS/man_model4you.Rnw
Torsten's comments
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2365 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: adf2266b1eaf897b98f98d22a1b1f79b7322b49b

\section*{Paper Author Roles and Affiliations}
1. PhD student; Biostatistics Department, Epidemiology, Biostatistics \&
Prevention Institute, University of Zurich; \new{Postdoc; Institute for Medical
Information Processing, Biometry, and Epidemiology, LMU Munich}

@@

\section*{Abstract}
Typical models estimating treatment effects assume that the treatment effect is
the same for all individuals.  Model-based recursive partitioning allows to
relax this assumption and to estimate stratified treatment effects (model-based
trees) or even personalised treatment effects (model-based forests).  With
model-based trees one can compute treatment effects for different strata of
individuals. The strata are found in a data driven fashion and depend on
characteristics of the individuals.  Model-based random forests allow for a
similarity estimation between individuals in terms of model parameters (e.g.\
intercept and treatment effect). The similarity measure can then be used to
estimate personalised models.  The R package \emph{model4you} implements these
stratified and personalised models \new{in the setting with two randomly
assigned treatments} with a focus on ease of use and interpretability so that
clinicians and other users can take the model they usually use for the
estimation of the average treatment effect and with a few lines of code get a
visualisation that is easy to understand and interpret.

@@

Currently the R package \emph{model4you} can handle only

@@

\paragraph{1. Model-based trees} identify subgroups where within the subgroups
the model parameters are similar and between groups the model parameters are
different. This is achieved by finding instabilities in the model parameters
with respect to a variable (characteristic) and recursively partitioning the
data into groups. If, for example the algorithm finds that men and women have
differing treatment effects, the data is partitioned into two subgroups.
Details on model-based trees in general can be found in
\citep{zeileis_model-based_2008} and for the special use case for stratified
treatment effect estimation in \citep{seibold_model-based_2016}.  Just a single
line of code lets the user compute a model-based tree in R:
<<eval=FALSE>>=
strat_models <- pmtree(base_model)
@
Note that \Sexpr{'pmtree()'} uses the data given in the call of the base model.
It automatically uses variables not used in the model formula (in the example
above \Sexpr{'response ~ treatment'}) as potential subgroup defining variables.
This can be edited using the \Sexpr{'zformula'} argument.

@@

\section*{Implementation and architecture}
The R package \emph{model4you} is focused on ease of use and interpretability.
Users can take a simple model that they know and understand as basis.
\new{
Models currently available for use are linear models (function \Sexpr{'lm()'}
in R), generalised linear models (\Sexpr{'glm()'}) and survival models
(\Sexpr{'survreg()'} and \Sexpr{'coxph()'} from the R package \emph{survival}).
Models are restricted to those with a single binary covariate (e.g. two
treatment options).  The user can simply plug the model object into
\Sexpr{'pmtree()'} or \Sexpr{'pmforest()'} depending on whether they want
subgroup wise or personalised models.
}

@@

The \emph{partykit} package provides the basis for functionalities in other
packages namely \emph{glmertree}, \emph{psychotree}, \emph{betareg} (all on
CRAN), \emph{trtf}, \emph{disttree}, \emph{lagsarlmtree} and \emph{palmtree}
(all available on R-Forge, publishing on CRAN planned).

@@

The software is intentionally written to make usage as simple as possible. The
most prominent use case are clinical trials where the assumption of an average
treatment effect for all patients is too strict and the efficacy of the
treatment depends on patient characteristics (e.g. gender, biomarkers, etc.).
For subgroup analyses (stratified treatment effects) model-based trees
(\Sexpr{'pmtree()'}) can be used; For personalised treatment effects
model-based forests (\Sexpr{'pmforest()'}) provide a way of estimating
similarity between patients and using this similarity measure to estimate
personalised models (\Sexpr{'pmodel()'}).  The target audience are people who
deal with heterogeneous treatment effects, such as medical researchers,
pharmaceutical companies or analysts in marketing (A-B testing).  In general
the software is useful to researchers dealing with scenarios where two
exposures are compared and responses of subjects possibly depend on other
variables.

>>>>

\section*{Paper Author Roles and Affiliations}
1. \new{Postdoc; Institute for Medical
Information Processing, Biometry, and Epidemiology, LMU Munich}

@@

\section*{Abstract}
Typical models estimating treatment effects assume that the treatment effect is
the same for all individuals.  Model-based recursive partitioning allows to
relax this assumption and to estimate stratified treatment effects (model-based
trees) or even personalised treatment effects (model-based forests).  With
model-based trees one can compute treatment effects for different strata of
individuals. The strata are found in a data driven fashion and depend on
characteristics of the individuals.  Model-based random forests allow for a
similarity estimation between individuals in terms of model parameters (e.g.~
intercept and treatment effect). The similarity measure can then be used to
estimate personalised models.  The R package \emph{model4you} implements these
stratified and personalised models \new{in the setting with two randomly
assigned treatments} with a focus on ease of use and interpretability so that
clinicians and other users can take the model they usually use for the
estimation of the average treatment effect and with a few lines of code get a
visualisation that is easy to understand and interpret.

@@

@@

\paragraph{1. Model-based trees} identify subgroups where within the subgroups
the model parameters are similar and between groups the model parameters are
different. This is achieved by finding instabilities in the model parameters
with respect to a variable (characteristic) and recursively partitioning the
data into groups. If, for example the algorithm finds that men and women have
differing treatment effects, the data are partitioned into two subgroups.
Details on model-based trees in general can be found in
\citep{zeileis_model-based_2008} and for the special use case for stratified
treatment effect estimation in \citep{seibold_model-based_2016}.  Just a single
line of code lets the user compute a model-based tree in R:
<<eval=FALSE>>=
strat_models <- pmtree(base_model)
@
Note that \Sexpr{'pmtree()'} uses the data given in the call of the base model.
It automatically uses variables not used in the model formula (in the example
above \Sexpr{'response ~ treatment'}) as potential subgroup defining variables.
This can be edited using the \Sexpr{'zformula'} argument.

@@

\section*{Implementation and architecture}
The R package \emph{model4you} is focused on ease of use and
interpretability.  Users can take a simple model that they know and
understand as basis.
\new{
Models currently available for use are linear models (function
\Sexpr{'lm()'} in R), generalised linear models (\Sexpr{'glm()'}) and
survival models (\Sexpr{'survreg()'} and \Sexpr{'coxph()'} from the R
package \emph{survival}).  Models are restricted to those with a
single binary covariate (e.g.~ two treatment options).  The user can
simply plug the model object into \Sexpr{'pmtree()'} or
\Sexpr{'pmforest()'} depending on whether they want subgroup wise or
personalised models.
}

@@

The \emph{partykit} package provides the basis for functionalities in
other packages namely \emph{glmertree}, \emph{psychotree},
\emph{betareg}, \emph{disttree}, \emph{lagsarlmtree}, \emph{palmtree}
(all on CRAN), and \emph{trtf} (currently avilable on R-Forge).

@@

The software is intentionally written to make usage as simple as possible. The
most prominent use case are clinical trials where the assumption of an average
treatment effect for all patients is too strict and the efficacy of the
treatment depends on patient characteristics (e.g.~ gender, biomarkers, etc.).
For subgroup analyses (stratified treatment effects) model-based trees
(\Sexpr{'pmtree()'}) can be used; For personalised treatment effects
model-based forests (\Sexpr{'pmforest()'}) provide a way of estimating
similarity between patients and using this similarity measure to estimate
personalised models (\Sexpr{'pmodel()'}).  The target audience are people who
deal with heterogeneous treatment effects, such as medical researchers,
pharmaceutical companies or analysts in marketing (A-B testing).  In general
the software is useful to researchers dealing with scenarios where two
exposures are compared and responses of subjects possibly depend on other
variables.
